B\'{e}zier $\bar{B}$ projection

### 2018-01-25 projection1120.tex
deleted all 'deleted' part

A global $\bar{B}$ method that utilizes a global $L^2$ projection results in a dense stiffness matrix. The B\'ezier $\bar{B}$ \replaced{methods}{method}, on the other hand, \replaced{produce}{produces} sparse stiffness matrices. However, the coupling of the local dual basis functions does increase the bandwidth slightly. This is illustrated in Figure~\ref{fig:stiffness_matrix}, which shows the structure of the stiffness matrix for the Timoshenko beam problem using the second order basis functions of maximal smoothness for a displacement-based method (Figure~\ref{fig:stiffness_matrix}a), global $\bar{B}$ method (Figure~\ref{fig:stiffness_matrix}b), symmetric B\'ezier $\bar{B}$ method (Figure~\ref{fig:stiffness_matrix}c) and non-symmetric \Bezier $\bar{B}$ method (Figure~\ref{fig:stiffness_matrix}d). The blank cells indicate zero terms in the matrix while colored cells show the location of nonzero terms.

@@

The beam has a rectangular cross-section and we use the following non-dimensional sectional and material parameters: length $l=10$, width $b=1$, thickness $t=0.01$, Young's modulus $E=10^9$, Poisson's ratio $\nu=0.3$, and a shear correction factor of $s=5/6$. A comparison of the normalized error in the $L^2$ norm for $w$, $\phi$, $M$ and $Q$ versus the number of degrees of freedom for polynomial degrees $p=1,2,3$ is shown in Figure \ref{fig:timoshenko_result}. Results computed using standard finite elements are labeled $Q_1$, $Q_2$, and $Q_3$. Results computed using a global $\bar{B}$ method are labeled $\mathcal{T}^{L^2}$, those computed with \replaced{the symmetric B\'ezier $\bar{B}$ method and the non-symmetric B\'ezier $\bar{B}$ method are labeled $S-\mathcal{T}^{P}$ and $NS-\mathcal{T}^{P}$, respectively. }{the B\'ezier $\bar{B}$ method are labeled $\mathcal{T}^{P}$}. As expected, the $Q_1$ results lock and the error remains virtually unchanged as the mesh is refined. Increasing the polynomial degree does reduce the locking effect, although the reduction is minor for the $Q_2$ results. \replaced{$\mathcal{T}^{L^2}$, $S-\mathcal{T}^{P}$ and $NS-\mathcal{T}^P$}{Both $\mathcal{T}^{L^2}$ and $\mathcal{T}^P$} are essentially locking free for all polynomial orders. The convergence rates for the $\bar{B}$ methods are at least $p+1$ for $w$, $p$ for $\phi$, $p-1$ for $M$, and $p-2$ for $Q$. These rates agree with those reported in \cite{kiendl_single-variable_2015} and are optimal. To reiterate, B\'{e}zier $\bar{B}$ methods produces the same convergence rates as the global $\bar{B}$ method \added{and the the error plots of $NS-\mathcal{T}^{P}$ for $\phi$, $M$ and $Q$ are identical to those of $\mathcal{T}^{L^2}$}.
\begin{figure}
    \centering
    \begin{subfigure}[b]{0.49\linewidth}        %% or \columnwidth
        \centering
        \includegraphics[width=\linewidth]{w-sin}
        \caption{}
        \vspace*{2mm}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\linewidth}        %% or \columnwidth
        \centering
        \includegraphics[width=\linewidth]{phi-sin}
        \caption{}
        \vspace*{2mm}
    \end{subfigure}
        \begin{subfigure}[b]{0.49\linewidth}        %% or \columnwidth
        \centering
        \includegraphics[width=\linewidth]{M-sin}
        \caption{}
        \vspace*{2mm}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\linewidth}        %% or \columnwidth
        \centering
        \includegraphics[width=\linewidth]{Q-sin}
        \caption{}
        \vspace*{2mm}
    \end{subfigure}
    \caption{Convergence studies for slenderness factor $l/t=10^{-3}$. Error in the $L^2$-norm for (a) displacement $w$, (b) rotation $\phi$, (c) bending moment $M$, and (d) shear force $Q$.}
    \label{fig:timoshenko_result}
\end{figure}

@@

\section{Volumetric locking: Nearly incompressible linear elasticity}
\label{sec:nearlyincompressible}
To demonstrate the use of \Bezier $\bar{B}$ \replaced{methods}{method} to alleviate volumetric locking effects we study the nearly incompressible elasticity problem in two dimensions. We start with the small strain tensor $\boldsymbol{\varepsilon}$, which is defined as the symmetric part of the displacement gradient, i.e., 
\begin{equation}
\varepsilon_{ij}=\dfrac{u_{i,j}+u_{j,i}}{2}.
\end{equation}
The stress tensor is related to the strain tensor through the generalized Hooke's law
\begin{equation}
\sigma_{ij}=c_{ijkl}\varepsilon_{kl}
\end{equation}
where, for isotropic elasticity, the elastic coefficients and stress tensor can be expressed in terms of the Lam\'{e} parameters $\lambda$ and $\mu$ as
\begin{align}
c_{ijkl} &=\mu(\delta_{ik}\delta_{jl}+\delta_{il}\delta_{jk})+\lambda\delta_{ij}\delta_{kl}\\
\sigma_{ij} &= \lambda\varepsilon_{kk}\delta_{ij} + 2\mu\varepsilon_{ij}.
\end{align}
The Lam\'e parameters $\lambda$ and $\mu$ are defined in terms of Young's modulus, $E$, and Poisson's ratio, $\nu$, as
\begin{align}
\lambda &=\dfrac{\nu{E}}{(1+\nu)(1-2\nu)}\\
\mu &=\dfrac{\nu{E}}{2(1+\nu)}.
\end{align}
we can write the strong form of linear elasticity as
\begin{align}
	\sigma_{ij,j}+f_i &= 0 \text{ in $\Omega$} \\
    u_i &= g_i\text{ on $\Gamma_{g_i}$} \\
    \sigma_{ij}n_j &= h_i\text{ on $\Gamma_{h_i}$}
\end{align}
where Dirichlet boundary conditions are applied on $\Gamma_{g_i}$, Neumann boundary conditions are applied on $\Gamma_{h_i}$, and the closure of the domain $\Omega$ is $\bar{\Omega}=\Omega\cup\Gamma_{g_i}\cup\Gamma_{h_i}$. \added{To desmonstrate the source of volumetric locking, we introduce the pressure term}
\begin{equation}
    p=-(\lambda+2\mu/3){\epsilon}_{ii}.
\end{equation}
If $\nu\rightarrow\dfrac{1}{2}$ then $\lambda$ becomes very large and \added{an additional constraint is applied to the volumetric strain as}
\begin{equation}
    {\epsilon}_{ii}=0.
\end{equation}
\subsection{Symmetric \Bezier $\bar{B}$ projection}
\subsubsection{The weak form}
The $\bar{B}$ approach for nearly incompressible linear elasticity splits the strain tensor $\varepsilon$ into volumetric and deviatoric strains and then replaces the volumetric strain with a projected strain. We begin with
\begin{equation}
\boldsymbol{\varepsilon}(\mathbf{u})=\boldsymbol{\varepsilon}^{vol}(\mathbf{u})+\boldsymbol{\varepsilon}^{dev}(\mathbf{u})
\end{equation}
where $\boldsymbol{\varepsilon}^{vol}=\dfrac{1}{3}\mathrm{tr}(\boldsymbol{\varepsilon})\mi$ is the volumetric strain and $\boldsymbol{\varepsilon}^{dev}=\boldsymbol{\varepsilon}-\dfrac{1}{3}\mathrm{tr}(\boldsymbol{\varepsilon})\mi$ is the deviatoric strain. The volumetric strain is then replaced by a projected volumetric strain $\bar{\boldsymbol{\varepsilon}}^{vol}$ and the new total strain becomes
\begin{equation}
\bar{\boldsymbol{\varepsilon}}=\bar{\boldsymbol{\varepsilon}}^{vol}+\boldsymbol{\varepsilon}^{dev}.
\end{equation}
The weak form can then be written as: find $\mathbf{u}\in{\mathcal{S}(\Omega)}$ such that for all $\mathbf{w}\in{\mathcal{V}(\Omega)}$
\begin{equation}
    \bar{a}\langle{\mathbf{w},\mathbf{u}}\rangle=\bar{l}\langle{\mathbf{w}}\rangle
\end{equation}
where
\begin{align}
    \bar{a}\langle{\mathbf{w},\mathbf{u}}\rangle&=\int_{\Omega}\bar{\epsilon}_{ij}(\mathbf{w})c_{ijkl}\bar{\epsilon}_{kl}(\mathbf{u})d\Omega, \\
    \bar{l}\langle{\mathbf{w}}\rangle&=\int_{\Omega}\mathbf{w}\cdot\mathbf{f}d\Omega+\int_{\Gamma_{h}}\mathbf{w}\cdot\mathbf{h}d\Gamma.
\end{align}

>>>>

A global $\bar{B}$ method that utilizes a global $L^2$ projection results in a dense stiffness matrix. The B\'ezier $\bar{B}$ {methods}, on the other hand, {produce} sparse stiffness matrices. However, the coupling of the local dual basis functions does increase the bandwidth slightly. This is illustrated in Figure~\ref{fig:stiffness_matrix}, which shows the structure of the stiffness matrix for the Timoshenko beam problem using the second order basis functions of maximal smoothness for a displacement-based method (Figure~\ref{fig:stiffness_matrix}a), global $\bar{B}$ method (Figure~\ref{fig:stiffness_matrix}b), symmetric B\'ezier $\bar{B}$ method (Figure~\ref{fig:stiffness_matrix}c) and non-symmetric \Bezier $\bar{B}$ method (Figure~\ref{fig:stiffness_matrix}d). The blank cells indicate zero terms in the matrix while colored cells show the location of nonzero terms.

@@

The beam has a rectangular cross-section and we use the following non-dimensional sectional and material parameters: length $l=10$, width $b=1$, thickness $t=0.01$, Young's modulus $E=10^9$, Poisson's ratio $\nu=0.3$, and a shear correction factor of $s=5/6$. A comparison of the normalized error in the $L^2$ norm for $w$, $\phi$, $M$ and $Q$ versus the number of degrees of freedom for polynomial degrees $p=1,2,3$ is shown in Figure \ref{fig:timoshenko_result}. Results computed using standard finite elements are labeled $Q_1$, $Q_2$, and $Q_3$. Results computed using a global $\bar{B}$ method are labeled $\mathcal{T}^{L^2}$, those computed with the symmetric B\'ezier $\bar{B}$ method and the non-symmetric B\'ezier $\bar{B}$ method are labeled $S-\mathcal{T}^{P}$ and $NS-\mathcal{T}^{P}$, respectively. As expected, the $Q_1$ results lock and the error remains virtually unchanged as the mesh is refined. Increasing the polynomial degree does reduce the locking effect, although the reduction is minor for the $Q_2$ results. $\mathcal{T}^{L^2}$, $S-\mathcal{T}^{P}$ and $NS-\mathcal{T}^P$ are essentially locking free for all polynomial orders. The convergence rates for the $\bar{B}$ methods are at least $p+1$ for $w$, $p$ for $\phi$, $p-1$ for $M$, and $p-2$ for $Q$. These rates agree with those reported in \cite{kiendl_single-variable_2015} and are optimal. To reiterate, B\'{e}zier $\bar{B}$ methods produces the same convergence rates as the global $\bar{B}$ method \added{and the the error plots of $NS-\mathcal{T}^{P}$ for $\phi$, $M$ and $Q$ are identical to those of $\mathcal{T}^{L^2}$}.
\begin{figure}
    \centering
    \begin{subfigure}[b]{0.49\linewidth}        %% or \columnwidth
        \centering
        \includegraphics[width=\linewidth]{w-sin}
        \caption{}
        \vspace*{2mm}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\linewidth}        %% or \columnwidth
        \centering
        \includegraphics[width=\linewidth]{phi-sin}
        \caption{}
        \vspace*{2mm}
    \end{subfigure}
        \begin{subfigure}[b]{0.49\linewidth}        %% or \columnwidth
        \centering
        \includegraphics[width=\linewidth]{M-sin}
        \caption{}
        \vspace*{2mm}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\linewidth}        %% or \columnwidth
        \centering
        \includegraphics[width=\linewidth]{Q-sin}
        \caption{}
        \vspace*{2mm}
    \end{subfigure}
    \caption{Convergence studies for slenderness factor $l/t=10^{-3}$. Error in the $L^2$-norm for (a) displacement $w$, (b) rotation $\phi$, (c) bending moment $M$, and (d) shear force $Q$.}
    \label{fig:timoshenko_result}
\end{figure}

@@

\section{Volumetric locking: Nearly incompressible linear elasticity}
\label{sec:nearlyincompressible}
To demonstrate the use of \Bezier $\bar{B}$ {methods} to alleviate volumetric locking effects we study the nearly incompressible elasticity problem in two dimensions. We start with the small strain tensor $\boldsymbol{\varepsilon}$, which is defined as the symmetric part of the displacement gradient, i.e., 
\begin{equation}
\varepsilon_{ij}=\dfrac{u_{i,j}+u_{j,i}}{2}.
\end{equation}
The stress tensor is related to the strain tensor through the generalized Hooke's law
\begin{equation}
\sigma_{ij}=c_{ijkl}\varepsilon_{kl}
\end{equation}
where, for isotropic elasticity, the elastic coefficients and stress tensor can be expressed in terms of the Lam\'{e} parameters $\lambda$ and $\mu$ as
\begin{align}
c_{ijkl} &=\mu(\delta_{ik}\delta_{jl}+\delta_{il}\delta_{jk})+\lambda\delta_{ij}\delta_{kl}\\
\sigma_{ij} &= \lambda\varepsilon_{kk}\delta_{ij} + 2\mu\varepsilon_{ij}.
\end{align}
The Lam\'e parameters $\lambda$ and $\mu$ are defined in terms of Young's modulus, $E$, and Poisson's ratio, $\nu$, as
\begin{align}
\lambda &=\dfrac{\nu{E}}{(1+\nu)(1-2\nu)}\\
\mu &=\dfrac{\nu{E}}{2(1+\nu)}.
\end{align}
we can write the strong form of linear elasticity as
\begin{align}
	\sigma_{ij,j}+f_i &= 0 \text{ in $\Omega$} \\
    u_i &= g_i\text{ on $\Gamma_{g_i}$} \\
    \sigma_{ij}n_j &= h_i\text{ on $\Gamma_{h_i}$}
\end{align}
where Dirichlet boundary conditions are applied on $\Gamma_{g_i}$, Neumann boundary conditions are applied on $\Gamma_{h_i}$, and the closure of the domain $\Omega$ is $\bar{\Omega}=\Omega\cup\Gamma_{g_i}\cup\Gamma_{h_i}$. \added{To desmonstrate the source of volumetric locking, we introduce the pressure term}
\begin{equation}
    p=-(\lambda+2\mu/3){\epsilon}_{ii}.
\end{equation}
If $\nu\rightarrow\dfrac{1}{2}$ then $\lambda$ becomes very large and \added{an additional constraint is applied to the volumetric strain as}
\begin{equation}
    {\epsilon}_{ii}=0.
\end{equation}
\subsection{Symmetric \Bezier $\bar{B}$ projection}
\subsubsection{The weak form}
The $\bar{B}$ approach for nearly incompressible linear elasticity splits the strain tensor $\varepsilon$ into volumetric and deviatoric strains and then replaces the volumetric strain with a projected strain. We begin with
\begin{equation}
\boldsymbol{\varepsilon}(\mathbf{u})=\boldsymbol{\varepsilon}^{vol}(\mathbf{u})+\boldsymbol{\varepsilon}^{dev}(\mathbf{u})
\end{equation}
where $\boldsymbol{\varepsilon}^{vol}=\dfrac{1}{3}\mathrm{tr}(\boldsymbol{\varepsilon})\mi$ is the volumetric strain and $\boldsymbol{\varepsilon}^{dev}=\boldsymbol{\varepsilon}-\dfrac{1}{3}\mathrm{tr}(\boldsymbol{\varepsilon})\mi$ is the deviatoric strain. The volumetric strain is then replaced by a projected volumetric strain $\bar{\boldsymbol{\varepsilon}}^{vol}$ and the new total strain becomes
\begin{equation}
\bar{\boldsymbol{\varepsilon}}=\bar{\boldsymbol{\varepsilon}}^{vol}+\boldsymbol{\varepsilon}^{dev}.
\end{equation}
The weak form can then be written as: find $\mathbf{u}\in{\mathcal{S}(\Omega)}$ such that for all $\mathbf{w}\in{\mathcal{V}(\Omega)}$
\begin{equation}
    \bar{a}\langle{\mathbf{w},\mathbf{u}}\rangle=\bar{l}\langle{\mathbf{w}}\rangle
\end{equation}
where
\begin{align}
    \bar{a}\langle{\mathbf{w},\mathbf{u}}\rangle&=\int_{\Omega}\bar{\epsilon}_{ij}(\mathbf{w})c_{ijkl}\bar{\epsilon}_{kl}(\mathbf{u})d\Omega, \\
    \bar{l}\langle{\mathbf{w}}\rangle&=\int_{\Omega}\mathbf{w}\cdot\mathbf{f}d\Omega+\int_{\Gamma_{h}}\mathbf{w}\cdot\mathbf{h}d\Gamma.
\end{align}
